---
title: "Effect of character correlation"
author: "Thomas Guillerme"
date: "`r Sys.Date()`"
output:
  html_document:
    fig_width: 12
    fig_height: 6
---

Effect of character correlation
===============


Loading the functions and the data
-----------------------------------

Getting the functions and the data

```{r, eval = TRUE, message = FALSE}
library(dispRity)
source("functions.R") ; load.functions(test = FALSE)
load("../Data/NTS/matrix.nts.Rda")
source("../Data/load.data.R")
```


# Results (plots)


Comparison to 'best' tree (RF normalised distance)

```{r, fig.width=12, fig.height=12}
op <- par(mfrow=c(3,1), mar = c(2, 4, 1, 2))
plot.results.single(t25_list, metric = 2, what = "best",
    legend = TRUE, NTS = TRUE, xlab = "", ylab = "RF for 25 taxa",
    axislab = FALSE)
plot.results.single(t75_list, metric = 2, what = "best", ylab = "RF for 75 taxa",
    xlab = "", axislab = FALSE, NTS = TRUE)
plot.results.single(t150_list, metric = 2, what = "best",
    ylab = "RF for 150 taxa", NTS = TRUE)
par(op)
```

<!-- For generating the figure for the paper (without the legend) -->

```{r, eval = TRUE, echo = FALSE, print = FALSE, result = 'hide'}
pdf("../Writing/Figures/RF_results_best.pdf",width = 8,height = 8)
op <- par(mfrow=c(3,1), mar = c(2, 4, 1, 2))
plot.results.single(t25_list, metric = 2, what = "best",
    legend = FALSE, NTS = TRUE, xlab = "", ylab = "RF for 25 taxa",
    axislab = FALSE)
plot.results.single(t75_list, metric = 2, what = "best", ylab = "RF for 75 taxa",
    xlab = "", axislab = FALSE, NTS = TRUE)
plot.results.single(t150_list, metric = 2, what = "best",
    ylab = "RF for 150 taxa", NTS = TRUE)
par(op)
# dev.off()
```


Comparison to 'best' tree (Triplets normalised distance)

```{r, fig.width=12, fig.height=12}
op <- par(mfrow=c(3,1), mar = c(2, 4, 1, 2))
plot.results.single(t25_list, metric = 4, what = "best",
    legend = TRUE, NTS = TRUE, xlab = "", ylab = "Tr for 25 taxa",
    axislab = FALSE)
plot.results.single(t75_list, metric = 4, what = "best", ylab = "Tr for 75 taxa",
    xlab = "", axislab = FALSE, NTS = TRUE)
plot.results.single(t150_list, metric = 4, what = "best",
    ylab = "Tr for 150 taxa", NTS = TRUE)
par(op)
```


<!-- For generating the figure for the paper (without the legend) -->

```{r, eval = TRUE, echo = FALSE, print = FALSE, result = 'hide'}
pdf("../Writing/Figures/Tr_results_best.pdf", width = 8, height = 8)
op <- par(mfrow=c(3,1), mar = c(2, 4, 1, 2))
plot.results.single(t25_list, metric = 4, what = "best",
    legend = FALSE, NTS = TRUE, xlab = "", ylab = "Tr for 25 taxa",
    axislab = FALSE)
plot.results.single(t75_list, metric = 4, what = "best", ylab = "Tr for 75 taxa",
    xlab = "", axislab = FALSE, NTS = TRUE)
plot.results.single(t150_list, metric = 4, what = "best",
    ylab = "Tr for 150 taxa", NTS = TRUE)
par(op)
# dev.off()
```


# Results supplementary

## Effect of character correlation compared to the null ("random") trees

Comparison to 'null' tree (RF normalised distance)

```{r, fig.width=12, fig.height=12}
op <- par(mfrow=c(3,1), mar = c(2, 4, 1, 2))
plot.results.single(t25_list, metric = 2, what = "null",
    legend = TRUE, NTS = TRUE, xlab = "", ylab = "RF for 25 taxa",
    axislab = FALSE)
plot.results.single(t75_list, metric = 2, what = "null", ylab = "RF for 75 taxa",
    xlab = "", axislab = FALSE, NTS = TRUE)
plot.results.single(t150_list, metric = 2, what = "null",
    ylab = "RF for 150 taxa", NTS = TRUE)
par(op)
```

<!-- For generating the figure for the paper (without the legend) -->

```{r, eval = TRUE, echo = FALSE, print = FALSE, result = 'hide'}
pdf("../Writing/Figures/RF_results_null.pdf", width = 8, height = 8)
op <- par(mfrow=c(3,1), mar = c(2, 4, 1, 2))
plot.results.single(t25_list, metric = 2, what = "null",
    legend = TRUE, NTS = TRUE, xlab = "", ylab = "RF for 25 taxa",
    axislab = FALSE)
plot.results.single(t75_list, metric = 2, what = "null", ylab = "RF for 75 taxa",
    xlab = "", axislab = FALSE, NTS = TRUE)
plot.results.single(t150_list, metric = 2, what = "null",
    ylab = "RF for 150 taxa", NTS = TRUE)
par(op)
# dev.off()
```

Comparison to 'null' tree (Triplets normalised distance)

```{r, fig.width=12, fig.height=12}
op <- par(mfrow=c(3,1), mar = c(2, 4, 1, 2))
plot.results.single(t25_list, metric = 4, what = "null",
    legend = TRUE, NTS = TRUE, xlab = "", ylab = "Tr for 25 taxa",
    axislab = FALSE)
plot.results.single(t75_list, metric = 4, what = "null", ylab = "Tr for 75 taxa",
    xlab = "", axislab = FALSE, NTS = TRUE)
plot.results.single(t150_list, metric = 4, what = "null",
    ylab = "Tr for 150 taxa", NTS = TRUE)
par(op)
```

<!-- For generating the figure for the paper (without the legend) -->

```{r, eval = TRUE, echo = FALSE, print = FALSE, result = 'hide'}
pdf("../Writing/Figures/Tr_results_null.pdf", width = 8, height = 8)
op <- par(mfrow=c(3,1), mar = c(2, 4, 1, 2))
plot.results.single(t25_list, metric = 4, what = "null",
    legend = FALSE, NTS = TRUE, xlab = "", ylab = "Tr for 25 taxa",
    axislab = FALSE)
plot.results.single(t75_list, metric = 4, what = "null", ylab = "Tr for 75 taxa",
    xlab = "", axislab = FALSE, NTS = TRUE)
plot.results.single(t150_list, metric = 4, what = "null",
    ylab = "Tr for 150 taxa", NTS = TRUE)
par(op)
# dev.off()
```

## Ability to recover the "true" tree

How different are the "normal" and "random" trees from the true trees

### 25 taxa

```{r, fig.width=12, fig.height=6}
## Comparisons to the true tree (RF)
plot.results.single(t25_list, metric = 2, what = "true",
    main = "comparison to 'true' tree (25 taxa)", legend = TRUE, NTS = TRUE)
```

```{r, fig.width=12, fig.height=6}
## Comparisons to the true tree (Triplets)
plot.results.single(t25_list, metric = 4, what = "true",
    main = "comparison to 'true' tree (25 taxa)", legend = TRUE, NTS = TRUE)
```

### 75 taxa

```{r, fig.width=12, fig.height=6}
## Comparisons to the true tree (RF)
plot.results.single(t75_list, metric = 2, what = "true",
    main = "comparison to 'true' tree (75 taxa)", legend = TRUE, NTS = TRUE)
```

```{r, fig.width=12, fig.height=6}
## Comparisons to the true tree (Triplets)
plot.results.single(t75_list, metric = 4, what = "true",
    main = "comparison to 'true' tree (75 taxa)", legend = TRUE, NTS = TRUE)
```

### 150 taxa

```{r, fig.width=12, fig.height=6}
## Comparisons to the true tree (RF)
plot.results.single(t150_list, metric = 2, what = "true",
    main = "comparison to 'true' tree (150 taxa)", legend = TRUE, NTS = TRUE)
```

```{r, fig.width=12, fig.height=6}
## Comparisons to the true tree (Triplets)
plot.results.single(t150_list, metric = 4, what = "true",
    main = "comparison to 'true' tree (150 taxa)", legend = TRUE, NTS = TRUE)
```



## Resolution

How well resolved are the trees?
These graphs represent the number of nodes (i.e. the tree resolution) in the consensus trees for 25, 75 or 50 taxa and 100, 350 or 1000 characters (c).
The solid and dashed lines are respectively the 50% and 95% confidence intervals.
The black dot is the median.

```{r, fig.width=12, fig.height=6}
## Nodes for the trees with 25 taxa
plot.tree.resolution("../Data/Trees_out/Consensus_trees/", ntaxa = 25, col = c("darkgrey", "red", "green3", "blue", "grey", "orange", "lightgreen", "lightblue"))
```

```{r, fig.width=12, fig.height=6}
## Nodes for the trees with 75 taxa
plot.tree.resolution("../Data/Trees_out/Consensus_trees/", ntaxa = 75, col = c("darkgrey", "red", "green3", "blue", "grey", "orange", "lightgreen", "lightblue"))
```

```{r, fig.width=12, fig.height=6}
## Nodes for the trees with 150 taxa
plot.tree.resolution("../Data/Trees_out/Consensus_trees/", ntaxa = 150, col = c("darkgrey", "red", "green3", "blue", "grey", "orange", "lightgreen", "lightblue"))
```