---
title: "Check matrix correlations"
author: "Thomas Guillerme"
date: "`r Sys.Date()`"
output:
  html_document:
    fig_width: 12
    fig_height: 6
---

# Checking the number of character replaced in each matrix


```{r}
library(ape)
## Setting the path
PATH <- "../Data/Simulations/Matrices/"
```

The following function counts, _a posteriori_ the number of characters that where replaced in each matrix (by comparing them to the normal matrix)

```{r}
## Getting the proportion of replaced characters
get.proportion.replace <- function(matrix, path) {
    ## Getting one matrix in a comparable format
    get.one.matrix <- function(path, matrix, type) {
        return(unlist(apply(do.call(rbind, read.nexus.data(paste0(path, matrix, "_", type, ".nex"))), 2, list), recursive = FALSE))
    }

    ## Getting all the matrices
    norm <- get.one.matrix(path, matrix, "norm")
    maxi <- get.one.matrix(path, matrix, "maxi")
    mini <- get.one.matrix(path, matrix, "mini")
    rand <- get.one.matrix(path, matrix, "rand")


    ## Comparing all the matrices
    n <- length(norm)
    maxi <- length(which(!mapply(function(X, Y) all(X == Y), norm, maxi)))/n
    mini <- length(which(!mapply(function(X, Y) all(X == Y), norm, mini)))/n
    rand <- length(which(!mapply(function(X, Y) all(X == Y), norm, rand)))/n

    return(c("maxi" = maxi, "mini" = mini, "rand" = rand))
}
```

We can then apply it to the list of matrices

```{r}
## Creating the placeholder list for the results
replacements <- list(list(), list(), list())
## getting the values to loop through
taxa_val <- c(25, 75, 150)
character_val <- c(100, 350, 1000)
for(taxa in 1:3) {
    for(character in 1:3) {
        ## Get the taxa names
        names <- unique(gsub("_[a-z]*.nex", "",
          list.files(PATH, pattern = paste0(taxa_val[taxa], "t_", character_val[character], "c_"))))
        ## Get the proportions of replaced characters
        replacements[[taxa]][[character]] <- lapply(as.list(names), get.proportion.replace, PATH)
    }
}

## Combining the results
fun.maxi <- function(list) {
    return(unlist(lapply(list, function(X) return(X[[1]]))))
}
fun.mini <- function(list) {
    return(unlist(lapply(list, function(X) return(X[[2]]))))
}
fun.rand <- function(list) {
    return(unlist(lapply(list, function(X) return(X[[3]]))))
}
results_maxi <- lapply(replacements, lapply, fun.maxi)
results_mini <- lapply(replacements, lapply, fun.mini)
results_rand <- lapply(replacements, lapply, fun.rand)

## Naming the lists
c("25t", "75t", "100t") -> names(results_maxi) -> names(results_mini) -> names(results_rand)
c("100c", "350c", "1000c") -> names(results_maxi[[1]]) ->
                              names(results_maxi[[2]]) ->
                              names(results_maxi[[3]]) ->
                              names(results_mini[[1]]) ->
                              names(results_mini[[2]]) ->
                              names(results_mini[[3]]) ->
                              names(results_rand[[1]]) ->
                              names(results_rand[[2]]) ->
                              names(results_rand[[3]])
```

We can then represent the results using box plots:

```{r, fig.height = 6, fig.width = }

boxplot(results_rand[[3]], xaxis = "n")





plot.results.single <- function(taxa_list, metric, what, ylim, legend = FALSE, NTS = TRUE, ylab, xlab, axislab, ...) {
    
    ## Graphic parameters
    ## Colours
    cols <- c("red", "orange", "green3", "lightgreen", "blue", "lightblue")
    
    ## Y lab
    if(missing(ylab)) {
        if(metric == 1) ylab <- "Similarity (Matching Cluster)"
        if(metric == 2) ylab <- "Similarity (Robinson-Foulds)"
        if(metric == 3) ylab <- "Similarity (Nodal Split)"
        if(metric == 4) ylab <- "Similarity (Triplets)"
    }

    ## Y lim
    get.min.list <- function(X, metric) {
        return(min(unlist(X[,metric])))
    }

    min <- min(unlist(lapply(unlist(unlist(unlist(taxa_list, recursive = FALSE), recursive = FALSE), recursive = FALSE), get.min.list, metric)))
    if(min < 0) {
        ylim <- c(min,1)
    } else {
        ylim <- c(0,1)
    }

    ## X lab
    if(missing(xlab)) {
        xlab <- "Characters"
    }

    ## axis lab
    if(missing(axislab)) {
        axislab <- c("100", "350", "1000")
    }

    ## Horizontal line
    line <- 0
    

    if(what == "null") {
        boxplot(xlab = xlab, ylab = ylab, xaxt = "n", ylim = ylim, col = rep(cols, 3), las = 1, ...,
            ## c100
            unlist(taxa_list$c100$bayesian$rand$maxi[,metric]), unlist(taxa_list$c100$parsimony$rand$maxi[,metric]),
            unlist(taxa_list$c100$bayesian$rand$mini[,metric]), unlist(taxa_list$c100$parsimony$rand$mini[,metric]),
            unlist(taxa_list$c100$bayesian$rand$norm[,metric]), unlist(taxa_list$c100$parsimony$rand$norm[,metric]),
            ## c350
            unlist(taxa_list$c350$bayesian$rand$maxi[,metric]), unlist(taxa_list$c350$parsimony$rand$maxi[,metric]),
            unlist(taxa_list$c350$bayesian$rand$mini[,metric]), unlist(taxa_list$c350$parsimony$rand$mini[,metric]),
            unlist(taxa_list$c350$bayesian$rand$norm[,metric]), unlist(taxa_list$c350$parsimony$rand$norm[,metric]),
            ## c1000
            unlist(taxa_list$c1000$bayesian$rand$maxi[,metric]), unlist(taxa_list$c1000$parsimony$rand$maxi[,metric]),
            unlist(taxa_list$c1000$bayesian$rand$mini[,metric]), unlist(taxa_list$c1000$parsimony$rand$mini[,metric]),
            unlist(taxa_list$c1000$bayesian$rand$norm[,metric]), unlist(taxa_list$c1000$parsimony$rand$norm[,metric])
            )

        ## X axis
        axis(1, 1:18, labels = FALSE, tick = FALSE)
        axis(1, c(3.5, 9.5, 15.5), tick = FALSE, labels = axislab)
        ## Lines
        abline(v = 6.5) ; abline(v = 12.5)
        abline(h = line, col = "grey", lty = 2)
        ## Legend
        if(legend) {
            legend("bottomleft", col = cols, pch = 19, bty = "n", title = "Character differences:", cex = 0.8,
                legend = c("Maximised (Bayesian)", "Maximised (Parsimony)",
                           "Minimised (Bayesian)", "Minimised (Parsimony)",
                           "Normal (Bayesian)", "Normal (Parsimony)")
                )
        }
    }

    if(what == "best") {
        boxplot(xlab = xlab, ylab = ylab, xaxt = "n", ylim = ylim, col = rep(cols, 3), las = 1, ...,
            ## c100
            unlist(taxa_list$c100$bayesian$norm$maxi[,metric]), unlist(taxa_list$c100$parsimony$norm$maxi[,metric]),
            unlist(taxa_list$c100$bayesian$norm$mini[,metric]), unlist(taxa_list$c100$parsimony$norm$mini[,metric]),
            unlist(taxa_list$c100$bayesian$norm$rand[,metric]), unlist(taxa_list$c100$parsimony$norm$rand[,metric]),
            ## c350
            unlist(taxa_list$c350$bayesian$norm$maxi[,metric]), unlist(taxa_list$c350$parsimony$norm$maxi[,metric]),
            unlist(taxa_list$c350$bayesian$norm$mini[,metric]), unlist(taxa_list$c350$parsimony$norm$mini[,metric]),
            unlist(taxa_list$c350$bayesian$norm$rand[,metric]), unlist(taxa_list$c350$parsimony$norm$rand[,metric]),
            ## c1000
            unlist(taxa_list$c1000$bayesian$norm$maxi[,metric]), unlist(taxa_list$c1000$parsimony$norm$maxi[,metric]),
            unlist(taxa_list$c1000$bayesian$norm$mini[,metric]), unlist(taxa_list$c1000$parsimony$norm$mini[,metric]),
            unlist(taxa_list$c1000$bayesian$norm$rand[,metric]), unlist(taxa_list$c1000$parsimony$norm$rand[,metric])
            )

        ## X axis
        axis(1, 1:18, labels = FALSE, tick = FALSE)
        axis(1, c(3.5, 9.5, 15.5), tick = FALSE, labels = axislab)
        ## Lines
        abline(v = 6.5) ; abline(v = 12.5)
        abline(h = line, col = "grey", lty = 2)
        ## Legend
        if(legend) {
            legend("bottomleft", col = cols, pch = 19, bty = "n", title = "Character differences:", cex = 0.8,
                legend = c("Maximised (Bayesian)", "Maximised (Parsimony)",
                           "Minimised (Bayesian)", "Minimised (Parsimony)",
                           "Randomised (Bayesian)", "Randomised (Parsimony)")
                )
        }
    }

    if(what == "true") {
        boxplot(xlab = xlab, ylab = ylab, xaxt = "n", ylim = ylim, col = rep(cols[1:4], 3), las = 1, ...,
            ## c100
            unlist(taxa_list$c100$bayesian$norm$true[,metric]), unlist(taxa_list$c100$parsimony$norm$true[,metric]),
            unlist(taxa_list$c100$bayesian$rand$true[,metric]), unlist(taxa_list$c100$parsimony$rand$true[,metric]),
            ## c350
            unlist(taxa_list$c350$bayesian$norm$true[,metric]), unlist(taxa_list$c350$parsimony$norm$true[,metric]),
            unlist(taxa_list$c350$bayesian$rand$true[,metric]), unlist(taxa_list$c350$parsimony$rand$true[,metric]),
            ## c1000
            unlist(taxa_list$c1000$bayesian$norm$true[,metric]), unlist(taxa_list$c1000$parsimony$norm$true[,metric]),
            unlist(taxa_list$c1000$bayesian$rand$true[,metric]), unlist(taxa_list$c1000$parsimony$rand$true[,metric])
            )

        ## X axis
        axis(1, 1:12, labels = FALSE, tick = FALSE)
        axis(1, c(2.5, 6.5, 10.5), tick = FALSE, labels = axislab)
        ## Lines
        abline(v = 4.5) ; abline(v = 8.5)
        abline(h = line, col = "grey", lty = 2)
        ## Legend
        if(legend) {
            legend("bottomleft", col = cols[1:4], pch = 19, bty = "n", title = "Character differences:", cex = 0.8,
                legend = c("Normal (Bayesian)", "Normal (Parsimony)",
                           "Randomised (Bayesian)", "Randomised (Parsimony)")
                )
        }
    }

    if(what == "check") {
        boxplot(xlab = xlab, ylab = ylab, xaxt = "n", ylim = ylim, col = rep(cols[1:4], 3), las = 1, ...,
            ## c100
            unlist(taxa_list$c100$bayesian$norm$rand[,metric]), unlist(taxa_list$c100$parsimony$norm$rand[,metric]),
            unlist(taxa_list$c100$bayesian$rand$norm[,metric]), unlist(taxa_list$c100$parsimony$rand$norm[,metric]),
            ## c350
            unlist(taxa_list$c350$bayesian$norm$rand[,metric]), unlist(taxa_list$c350$parsimony$norm$rand[,metric]),
            unlist(taxa_list$c350$bayesian$rand$norm[,metric]), unlist(taxa_list$c350$parsimony$rand$norm[,metric]),
            ## c1000
            unlist(taxa_list$c1000$bayesian$norm$rand[,metric]), unlist(taxa_list$c1000$parsimony$norm$rand[,metric]),
            unlist(taxa_list$c1000$bayesian$rand$norm[,metric]), unlist(taxa_list$c1000$parsimony$rand$norm[,metric])
            )

        ## X axis
        axis(1, 1:12, labels = FALSE, tick = FALSE)
        axis(1, c(2.5, 6.5, 10.5), tick = FALSE, labels = axislab)
        ## Lines
        abline(v = 4.5) ; abline(v = 8.5)
        abline(h = line, col = "grey", lty = 2)
        ## Legend
        if(legend) {
            legend("bottomleft", col = cols[1:4], pch = 19, bty = "n", title = "Character differences:", cex = 0.8,
                legend = c("Normal (Bayesian)", "Normal (Parsimony)",
                           "Randomised (Bayesian)", "Randomised (Parsimony)")
                )
        }
    }
}

combine_results <- unlist(lapply(replacements[[1]][[1]], function(X) return(X[[1]])))

